---
title: "IFSpakietJKM"
author: "Katarzyna Piskala, Maja Szlenk, Jakub Sztachelski"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IFSpakietJKM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```

```{r loading package, echo=FALSE, warning=FALSE, message=FALSE}
library(IFSpakietJKM)
```

#Introduction 

The package creates an objects of class IFS (class defined in the package) and generates plots of these objects. Class IFS is defined as a list of functions (contractions) which is used to generate fractals.
    
Functionality of the package will be shown in this vignette.

#Our package contains some useful functions.

###check_if_f(f)
The function checks if its argument is a list of functions from $\mathbb{R}^2$ to $\mathbb{R}^2$. Otherwise it returns error. This function can be used by a user in his code but also it is used in the function createIFS described below.

###createIFS(f)
The function creates the IFS (Iterated Function System) object from its argument. An object of class IFS is the representation of a fractal. Our package allows to generate an image of this fractal using generic function plot.

###cool_colours(n,d,...)
Returns a palette of colours which allows to plot the fractal in gradually changing colours.
The function cool_colours creates a vector of colours to put as a col parameter in plot.IFS. The length of this vector is equal $n^d$, where $n$ is a number of functions making a fractal and $d$ denotes density of a fractal. The other arguments in cool_colors are parameters of the rainbow function (see ? rainbow).

###plot.IFS
The overweighted generic function that draws fractal from an IFS object to given density $d$ ($d = 8$ by default).

# A simple example is presented below.
### Creating a list of contractions.
To make a fractal, the user needs to create a list of functions from $\mathbb{R}^2$ to $\mathbb{R}^2$. An example below shows functions creating Sierpi?ski triangle. 


``` {r list of contractions}
sierpinski_1 <- function(x){
  c(x[1]/2,x[2]/2)
}

sierpinski_2 <- function(x){
  c(x[1]/2+0.5, x[2]/2)
}

sierpinski_3 <- function(x){
  c(x[1]/2+0.25, x[2]/2+0.4330127)
}

sierpinski_list <- list(sierpinski_1,sierpinski_2,sierpinski_3)
```

Our package contains four list creating fractals: Sierpinski triangle, Heighway dragon, Koch snowflake and Barnsley's fern.


### Creating an IFS object.
Having the list of functions, the user needs to create an IFS object by using IFSpakietJKM package.

``` {r IFS}
library(IFSpakietJKM)
sierpinski_triangle <- createIFS(sierpinski_list)

```

### Drawing an IFS object.
Now all the user has to do is draw an object.
``` {r draw}
plot(sierpinski_triangle)
```



#Additional examples:

### Density change ("d" parameter).
```{r smok}
plot(createIFS(dragon), d=13)
```

### Changing parameters in a plot.
```{r}
barnsley <- createIFS(fern)
plot(barnsley, d=6, col="green", pch=20)
```

### Rainbow colours as a "col" parameter.
```{r trojkat}
plot(createIFS(triangle), col=cool_colours(3,8))
```

### Parameters of a cool_colours function.
```{r blue}
blue <- cool_colours(3,8, start=1/2, end=5/7)
plot(createIFS(triangle), col=blue)
```

### Have fun!
