---
title: "IFSpremium"
author: "Zofia Grochulska, Michał Futrega"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### IFSpremium package
Our package creates plots of fractals that can be generated via Iterated Function System. It is based on the S3 class. Here we will briefly discuss the idea of an IFS fractal, show how the package should be used and how it is built and present some figures.

### A Few words about fractals
Fractals are amazing mathematical objects. You can think of them as figures that have a special property called "self - similarity". No matter how close you look at them - they look exactly the same. Fractals do not only look mind-blowing to a layman, but they also possess properties that are of mathematicians' interests.

A wide class of fractals (called IFS fractals) can be described in terms of contractions. These are functions that, intuitively, cause the function's image to become smaller and smaller. All you need to create such a fractal is a list of functions that generate the fractal.

### Usage
```{r include = FALSE}
library(IFSpremium)
```

1. Think of an IFS fractal you want to draw. Get the list of contractions that generate it. For example:
```{r include = TRUE}
myContractions <- list(
  function(x) {
    c((-3*x[1]+3*sqrt(3)+3)/((1+sqrt(3)-x[1])^2 + (x[2])^2) - (1+sqrt(3))/(2+sqrt(3)), 
    3*x[2]/((1+sqrt(3)-x[1])^2 + (x[2])^2))
  },
  
  function(x) {
    c(1/2*((3*x[1]-3*sqrt(3)-3-3*sqrt(3)*x[2])/((1+sqrt(3)-x[1])^2 + (x[2])^2) + (1+sqrt(3))/(2+sqrt(3))), 
    1/2*((-3*sqrt(3)*x[1]-3*x[2]+3*sqrt(3)+9)/((1+sqrt(3)-x[1])^2 + (x[2])^2) - (3+sqrt(3))/(2+sqrt(3))))
  },
  
  function(x) {
    c(1/2*((3*x[1]+3*sqrt(3)*x[2]-3*sqrt(3)-3)/((1+sqrt(3)-x[1])^2 + (x[2])^2) + (1+sqrt(3))/(2+sqrt(3))), 
    1/2*((3*sqrt(3)*x[1]-3*x[2]-3*sqrt(3)-9)/((1+sqrt(3)-x[1])^2 + (x[2])^2) + (3+sqrt(3))/(2+sqrt(3))))
  }
)
```

2. Define an object of IFS class via _createIFS()_. The function's only argument is the list of contractions.
```{r include = TRUE}
myFractal <- createIFS(myContractions)
```
You can also look up the IFS contractions included in the package (full list included below). The one above could be simply called by:
```{r include = TRUE}
myFractal <- createIFS(ApollonianGasket)
```

3. Plot the fractal. You need to pass the name of the IFS object and the number of iterations (the depth of plotting). If needed, you can specify the set of starting points. It is set to the \((0,0)\) point as default and in the case of the Apollonian Gasket we do not need anything more.

And now we are ready to use the _plot.IFS()_ function.

```{r fig.align = "center", fig.height = 5, fig.width = 5}
plot(myFractal, 4)
```
The fractal looks a bit pale, let us see how the resolution of the plot changes with respect to the depth of plotting.

```{r fig.align = "center", fig.height = 5, fig.width = 5}
plot(myFractal, 7)
```
It should be noticed that in the plot there are always as many colors as there are contractions in the contraction list of a fractal. One color represents one contraction so that we can see what points are generated via which function.

### Built-in fractals
In this package you will find lists of contractions for most popular fractals, e.g.:

* Apollonian Gasket
* Koch Curve
* Koch Snowflake
* Sierpinski Triangle
* Sierpinski Carpet
* Sierpinski Pentagon
* Pythagorean Tree

It is interesting to note that all aforementioned fractals except for Apollonian Gasket are simply affine maps. Whereas Apollonian Gasket is the limit point of iterated circle inversions.

### Some examples
To plot a nice Pythagorean Tree it is best to use a square as a set of starting points. Remember that the set of starting points should be passed as a list of two-element lists (the first is the x coordinate, the second - the y coordinate).
```{r include = TRUE}
points <- seq(0, 1, 1/10)
square <- list()
for (i in 1:11) 
  for (j in 1:11) 
    square <- append(square, list(c(points[i], points[j])))
```

```{r fig.align = "center", fig.height = 5, fig.width = 5}
myFractal <- createIFS(PythagoreanTree)
plot(myFractal, 6, startPts = square)
```
How would it look with one default point?

```{r fig.align = "center", fig.height = 5, fig.width = 5}
myFractal <- createIFS(PythagoreanTree)
plot(myFractal, 9)
```
Well, not good (and we had to increase the number of iterations so that it looks any good!)

Last but not least: Koch's Snowflake.
```{r fig.align = "center", fig.height = 5, fig.width = 5}
myFractal <- createIFS(KochSnowflake)
plot(myFractal, 6)
```

### How does the package work?
Two basic functions in the package are _createIFS()_ and _plot.IFS()_. We created also some helping functions _colorPoints()_ and _createPoints()_. If you are interested in the details of how the functions work have a look at the function's documentation. Here only an outline of the method will be discussed.

The _createIFS()_ function creates an object of IFS class so that the generic _plot()_ function can be called to draw the fractal. The aim of the  _colorPoints()_ points function is to ensure that the image of each contraction is painted in different color. The most tricky function is _createPoints()_, the one that returns a list of points of which the fractal consists. The list is created via two "for" loops, the outer one dedicated to iterating over the number of iterations, the inner iterates over possible contractions.

