---
title: "raport PISA 1/1"
author: Ilona Grabowicz, Anna Macioszek, Małgorzata Pawlak
geometry: left=2cm,right=2cm,top=2cm,bottom=2.5cm
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
#- \lhead{}
- \rfoot{\hyperlink{page.1}{HOME}}
- \hypersetup{colorlinks=true, linkcolor=[rgb]{0, 0.45, 0.55}}
- \setlength{\headsep}{1.5cm}
- \renewcommand{\footrulewidth}{0.4pt}
output:
  pdf_document: default
  word_document: default
abstract: Influence of mother's working situation on daughters' views on career benefits of learning mathematics and on their own math abilities. Based on data from PISA2012, s survey which tested 15-year-olds in reading, mathematics and science focussing on mathematics in 2012.
toc: yes
---


```{r setup, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# CONTEXT

[A new working paper](https://hbswk.hbs.edu/item/kids-benefit-from-having-a-working-mom) by Kathleen McGinn and her colleagues for Harvard Business School purports that working mothers are more likely to raise successful daughters.

> *"THERE ARE VERY FEW THINGS...THAT HAVE SUCH A CLEAR EFFECT ON GENDER INEQUALITY AS BEING RAISED BY A WORKING MOTHER."*

*Kathleen L. McGinn, the Cahners-Rabb Professor of Business Administration at Harvard Business School*

# ANALYSIS

We picked girls' responses to two specific questions and looked at it from a perspective of their mothers' working situation. Mother's working situation could be defined as one of the following: working full time (for pay), working part time (for pay), not working, but looking for a job or other (e.g. home duties).The two chosen questions were statements to which the girls needed to state to what extent they apply to them. Response scale had 4 following levels: strongly agree, agree, disagree, strongly disagree. First statement concerned the feeling about the usefullness of mathematics in their future career and read: "Learning mathematics is worthwhile for me because it will improve my career". The second was about the level of confidence the girls have in their own math abilities: "If I wanted to, I could do well in mathematics".   An interesting result was that the group who tend to have the most different results are girls whose mothers work part-time.


```{r data, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align="center", fig.height=3.8}



#install.packages(tidyr)
#install.packages(ggplot2)
#install.packages(stringr)
#install.packages(dplyr)
#install.packages(pander)
#install.packages(PISA2012lite)

library(tidyr)
library(ggplot2)
library(stringr)
library(dplyr)
library(pander)
library(PISA2012lite)


# function to process data
# select desired columns and remove NAs
prepare_data <- function(data, groups, values, filter_criteria = TRUE) {
data %>%
  filter(filter_criteria) %>%
  select(CNT, groups = groups, values = values) %>%
  na.omit %>%
  group_by(groups, values, CNT) %>%
  summarise(n = n()) %>%
  return()
}

# functions to calculate plot's parameters
calculate_bar_widths <- function(data) {
  data %>%
    ungroup() %>%
    select(-values) %>%
    group_by(groups, CNT) %>%
    summarise(total = sum(n, na.rm = T)) %>%
    spread(CNT, total) %>%
    return()
}

calculate_breaks_positions <- function(widths, spacing_fraction = 40) {
  tmp <- c()
  last_position = 0
  for (i in 1:length(widths)) {
    current = widths[i]
    previous = widths[i-1]
    if (i==1)
      previous = 0
    position = previous/2 + current/2 + last_position
    last_position = position
    tmp = c(tmp, position)
  }
  spacing <- sum(widths) / spacing_fraction
  breaks <- tmp + cumsum(rep(spacing, length(widths)))
  return(breaks)
}

prepare_labels <- function(data,
                           width_thresholds = c(0.09, 0.2),
                           wrap_widths = c(40,18,12)) {
  labels = unique(data$groups)
  bars_positions = data$bars_positions[seq(1, nrow(data),
                                           length(labels))]
  bars_positions = bars_positions / max(bars_positions)
  wrapped_labels = c()
  diffs = diff(bars_positions)
  diffs = c(diffs, max(bars_positions))
  for (i in 1:length(labels)) {
    if (diffs[i] <= width_thresholds[1])
      wrapped_labels[i] = str_wrap(labels[i],
                                   width = wrap_widths[1])
    if (diffs[i] > width_thresholds[1])
      wrapped_labels[i] = str_wrap(labels[i],
                                   width = wrap_widths[2])
    if (diffs[i] > width_thresholds[2])
      wrapped_labels[i] = str_wrap(labels[i], 
                                   width = wrap_widths[3])
  }
  return(wrapped_labels)
}

# plotting functions
make_plot <- function(data, country,
                      ...) {
  data_to_plot <- spread(data, CNT, n)
  widths <- calculate_bar_widths(data)
  widths <- unlist(widths[country])
  breaks_positions = calculate_breaks_positions(widths)
  data_to_plot$bars_positions = rep(breaks_positions, each = 4)
  labels <- prepare_labels(data_to_plot)
  draw_barplot(data_to_plot, country,
               widths, labels, breaks_positions,
               ...)
}

draw_barplot <- function(data_to_plot, country,
                         widths, labels, breaks_positions,
                         title, subtitle,
                         xlab, ylab,
                         color_values) {
  print(
    ggplot(data_to_plot,
           aes(x = bars_positions,
               y = data_to_plot[country],
               fill = values)) +
    geom_bar(width = rep(widths, each = 4),
             stat = "identity",
             position = "fill") + 
    ggtitle(title, subtitle = subtitle) +
    theme(plot.title = element_text(hjust = 0.5,
                                    size = 20,
                                    colour = "grey90",
                                    face = "bold"),
          plot.subtitle = element_text(hjust = 0.5,
                                       size = 11,
                                       face = "italic",
                                       family = "Times")) +
    scale_x_continuous(breaks = breaks_positions,
                       labels = labels) +
    theme(axis.text.x = element_text(hjust = 1,
                                     angle = 45,
                                     size = 7,
                                     colour = "grey30")) +
    theme(axis.text.y = element_text(size = 7,
                                     colour = "grey30")) +
    theme(axis.title.x = element_text(face = "bold.italic",
                                      size = 9,
                                      colour = "grey90")) +
    theme(axis.title.y = element_text(face = "bold.italic",
                                      size = 9,
                                      colour = "grey90")) +
    xlab(xlab) +
    ylab(ylab) +
    #theme(legend.title = element_text(face = "bold", size=9, colour = "grey30")) +
    theme(legend.text = element_text(size = 9,
                                     face = "italic",
                                     family = "Times",
                                     colour = "grey30")) +
    theme(legend.background = element_rect(fill = "grey95")) +
    scale_fill_manual(values = color_values) +
    theme(legend.title = element_blank())
  )
}

# making a page in with two barplots in a pdf output
draw_a_page <- function(data, country, ...) {

  arguments = list(...)
  
  cat("\\newpage")
  cat("\\vspace{40pt}")
  cat("\\begin{center}")
  cat("\\subsection{", country, "}")
  cat("\\end{center}")
  cat("\\vspace{20pt}")

  for (i in 1:2) {
    make_plot(data[[i]], country,
      title = arguments$titles[i],
      subtitle = arguments$subtitles[i],
      xlab = arguments$xlabs[i], ylab = arguments$ylabs[i],
      color_values = arguments$color_values[[i]])

      if(i == 1)
        cat("\\vspace{30pt}")
  }
}

make_two_plots_per_country <- function(data, groups, values, filter_criteria, ...) {
  data1 <- prepare_data(data, groups[1], values[1], filter_criteria[[1]])
  data2 <- prepare_data(data, groups[2], values[2], filter_criteria[[2]])
  countries = unique(data$CNT)
  for (country in countries) {
    draw_a_page(list("1" = data1, "2" = data2), country, ...)
  }
}

# finally, generating actual output

filter_criteria = (student2012$ST04Q01 == "Female")
color_values = c("#3F9C9F", "#55B9BC", "#EA7382", "#E4465B")

make_two_plots_per_country(data = student2012,
                           groups = rep("ST15Q01", 2),
                           values = c("ST29Q05", "ST43Q05"),
                           filter_criteria = list("1" = filter_criteria,
                                                  "2" = filter_criteria),
                           titles = c("MATH AND CAREER","MATH CONFIDENCE"),
                           subtitles = c(
                             "\n Learning mathematics is worthwhile for me \n because it will improve my career\n",
                             "\n If I wanted to, I could do well in mathematics\n"),
                           xlabs = rep("MOTHER'S OCCUPATION", 2),
                           ylabs = rep("FRACTION OF ANSWERS", 2),
                           color_values = list("1" = color_values,
                                               "2" = color_values))


```
