---
title: "IFSrainbow Package Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aesthetic specifications}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(ggplot2)
library(devtools)
library(fields)
library(caTools)
library(dplyr)
library(testthat)
load_all(".")
knitr::opts_chunk$set(fig.dpi = 96)
```

This vignette describes in a non-technical way how to create, modify and plot
fractals using `IFSrainbow` package. A nice interactive presentation of what this package is capable of 
can be found [here](https://lrajkowski.shinyapps.io/fractalmixer/).

## A bit of theory

*Iterated Functions System Fractal* (or *IFS fractal*) is defined by a finite
family of space transformations $\mathcal{F}=\{f_1,f_2,\ldots,f_k\}$ (typically *contractions*, 
i.e. "shrinking" functions). It is the invariant set of so called *Hutchinson's operator*,
which is the union of all possible finite compositions of functions in $\mathcal{F}$.
Examples of fractals that can be obtained in such way include famous *Sierpiński's Trangle*, 
*Koch's Snowflake* or *Heighway's Dragon*. See 
[this Wikipedia page](https://en.wikipedia.org/wiki/Iterated_function_system) for more details. 

## Fractal creation

As the IFS fractal is defined by the set of functions, `IFSpackage` introduces an S3 class `ifs` which
is no other than a list of functions. For example, in order to create *Heighway's Dragon* defined by the
following two functions
```{r}
f1 <- function(p){
  return( c(
    p[1]/2 + p[2]/2 - 1, 
    -p[1]/2 + p[2]/2
    ) )
}
f2 <- function(p){
  return( c(
    -p[1]/2 + p[2]/2,
    -(p[1]/2 + p[2]/2 - 1)
    ) )
}
```
one needs to use `createIFS` function. It accepts two "types" of arguments -- firstly it can be just a
sequence of functions or else a list of functions, so any of the two below would work
```{r}
my_heighway1 <- createIFS(f1, f2)
my_heighway2 <- createIFS( list(f1, f2) )
```
Of course, this can be performed for any number of functions.

## Plotting

Objects of class `ifs` can be plotted just by applying `plot` functions. 
Its default version use so called *chaos game* method of drawing the fractal,
in which random compositions of given functions are applied to a random set of points
to obtain an approximation of the "true" fractal.

The functions assumes three mandatory arguments:

+ `ifs`, which is the fractal in question
+ `n`, which is the number of points in the plot
+ `d`, which is the length of random compositions applied 

There are also optional arguments:

+ `probs`, which is the sequences of probabilities of choosing respective functions

Moreover, as we use the benefits of the `ggplot2` package, some additional graphical 
parameters can be passed to the `ggplot` function inside. Therefore the following
can be used to obtain a nice approximation of the Heighway Dragon
```{r}
plot( my_heighway1, 1000, 20 )
```

If we do not like colours (but why not?), prefer smaller points and wish that one of
the functions is used more often than the other, we could run
```{r}

plot( my_heighway1, 1000, 20,
      probs=c(.7,.3), color="black", size=.8 )
```

## Creating gifs

Gif animations can be created by function createIFSanimation. This function takes 6 arguments.

+ `startPoint`, the point from which we start to create IFS
+ `depth`, the number of gifs frames
+ `fractal`, fractal, that will be drawn 
+ `resolution`, because successive gifs frames are represented as matrices, we have to assume a resolution that means accuracy when coloring the corresponding matrix fields
+ `filename`, the name of the file being saved
+ `scale`, because the size of the gif depends on the fractal, depth and resolution, so the scale enlarges the figure proportionally

Eight first frames of the Sierpinski triangle
```{r}
fractal <- sierpinski
startPoint = c(0,0)
depth = 6
resolution = 2
scale = 1


for(l in 1:depth){
        IFS <- generateIFSgif(startPoint,l,fractal)
        f <- round(IFS,2)
        xmin <- min(f[,1])
        xmax <- max(f[,1])
        ymin <- min(f[,2])
        ymax <- max(f[,2])
        diamy <- (ymax-ymin)*10^2+2
        diamx <- (xmax-xmin)*10^2+2
    }

    DM <- array(0,c(diamx*scale,diamy*scale,depth))

    for(l in 1:depth){
        IFS <- generateIFSgif(startPoint,l,fractal)
        f <- round(IFS,resolution)
        f <- dplyr::distinct(f,x,y)

        for(i in 1:nrow(f)){

            x<-(round((f[i,1]-min(f[,1]))*10^resolution)+1)*scale
            y<-(round((f[i,2]-min(f[,2]))*10^resolution)+1)*scale
            DM[x,y,l]<-1
        }
    }

    image <- (DM*120)+50

    for(i in 1:depth){
        image(image[,,i], col=tim.colors(256))
    }
```
