---
title: "Usage"
author: "Zuzanna Baran"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package is designed to be easy way of generating fractals. We will go through
couple simple steps to generate fractal for Koch's snowflake.


## Creating functions

First we need to create functions which will be used for fractal. In case of Koch's snowflake this are the functions.

```{r, fig.show='hold', results='hide'}
koch_f1 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/2 * x - sqrt(3)/6 * y, sqrt(3)/6 * x + 1/2 * y)
}

koch_f2 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/3 * x + 1/sqrt(3), 1/3 * y + 1/3)
}

koch_f3 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/3 * x, 1/3 * y + 2/3)
}

koch_f4 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/3 * x - 1/sqrt(3), 1/3 * y + 1/3)
}

koch_f5 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/3 * x - 1/sqrt(3), 1/3 * y - 1/3)
}

koch_f6 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/3 * x, 1/3 * y - 2/3)
}

koch_f7 <- function(point) {
  x<- point[1]
  y<- point[2]
  c(1/3 * x + 1/sqrt(3), 1/3 * y - 1/3)
}
```

## Creating IFS object

When we have functions next step is to create IFS object we can do it calling `createIFS` function from `IFSZuzanna` package.

```{r, fig.show='hold', results='hide', message = FALSE, warning = FALSE}
library(IFSZuzanna)

kochSnowflake <- createIFS(list(koch_f1, koch_f2, koch_f3, koch_f4, koch_f5, koch_f6, koch_f7))
```


## Ploting IFS object

Once we have IFS object now last step is to plot it. We can choose depth of recursion and color in which we want to draw a fractal. Default color is black.

```{r, fig.show='hold'}
plot(kochSnowflake, 3, 'blue')
plot(kochSnowflake, 4, 'red')
plot(kochSnowflake, 5, 'black')
plot(kochSnowflake,6, 'grey')

```

